### 2022-03-24 paper/output/document/thesis_5_discussion.tex
i hate don knuth
full send

First, I explore how my system can be used in practice to guide AI agents,
help academic philosophers make philosophical progress, and augment the everyday practical reasoning 
that we all perform as we navigate the world. I discuss these issues in Sections \ref{AIethics}, 
\ref{computationalethics}, and \ref{ordinarypeople} respectively and outline the further work
necessary to bring each of these applications to life. Much of this work has to do with how the input
maxim given to my system is formulated and the role of common sense or prior knowledge.

In Section \ref{amapossible}, I discuss theoretical objections to automated ethics, derived from the
common philosophical intuition that there is no algorithm for ethics. I explore concrete versions of 
this objection and make a theoretical argument for the possibility of the kind of automated ethics 
I implement in this thesis. Finally, in Section \ref{relatedwork}, I situate my project among related work.%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsection{Automated Moral Agents in Practice \label{AIethics}%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
In this section, I outline the additional components necessary for my system
to guide AI agents in practice. As it stands, my system is a categorical imperative library that can
evaluate the moral status of a maxim. My project potentially 
serves as one component of an ``ethics engine'' that an AI agent could use to make ethical decisions.
For example, when an AI agent faces a moral dilemma in some internal representation, it could pass it 
to an input parser that translates the dilemma into a maxim in my logic. An output parser could then 
translate my system's output into a prescription for action that the AI agent could act on. \begin{figure}
\includegraphics[scale=0.43]{AI_engine.png}
\caption{An example of an ethics engine for an artificial agent. This ethics engine passes a moral dilemma 
through an input parser, applies the automated categorical imperative test, and finally processes the 
output using an output parser, producing a prescription for action. I contribute the automated categorical 
imperative component.} \label{fig:AIengine}
\end{figure} Figure \ref{fig:AIengine} depicts the workflow of this example ethics engine. In order 
for my system to guide an AI agent using this workflow, future work must develop such input and output parsers.

One of the biggest challenges for an ethics engine is the development of an input parser. An input parser 
for my implementation of automated Kantian ethics must translate a complex real-world situation into a flat 
circumstances, act, goal triple.
This requires that the input parser determine which circumstances are morally relevant
to a maxim, a controversial judgement. As introduced in Section \ref{whatisamaxim},
there is robust debate on the circumstances that should be considered when formulating a maxim, 
inspired by a common criticism of Kantian ethics called the tailoring objection. Recall that the 
tailoring objection is the worry that arbitrarily specific 
circumstances render any maxim universalizable. For example, consider the maxim ``When my name is Lavanya Singh 
and I am wearing a purple shirt and it is November 26th, I will lie in order to get some easy cash.'' 
Even if this maxim is willed universally, the circumstances are so 
specific that lying will not become the general mechanism for getting easy cash, so the lender will 
believe my lie and the maxim will remain effective. By tailoring the circumstances, any maxim can 
evade universalization.

The Kantian response to this criticism is to require that the circumstances included in the formulation
of the maxim be morally relevant. In the example above, my purple shirt and the date have no bearing on 
the moral status of lying. On the other hand, consider the maxim, ``When I am unemployed, I will murder
someone in order to take their job.'' The circumstance ``when I am unemployed'' clearly has some moral
relevance to the murder in question; it speaks to the motivation for the murder.

@@

Translating everyday situations into appropriate maxims is the bulk of the reasoning that a Kantian 
human being does when making decisions, so it is unsurprising that formulating a maxim 
is one of the biggest obstacles to using my categorical imperative library
in an AI ethics engine. One solution is for a human being to perform the role of the input
parser by supervising the operation of an AI agent. When the agent stumbles onto
an ethical dilemma, the human could take over, formulate
the right question, and feed it into the categorical imperative library to see what action the categorical 
imperative would prescribe. Alternatively, AI developers can identify expected ethical dilemmas 
that the machine may face in advance and hardcode their own judgements for these dilemmas. 
For proponents of the ``human-in-the-loop'' model of AI ethics, in which
ethical AI requires that humans guide machines, this kind of human involvement may be a feature \citep{loop}.
Both of these solutions imply that the outcome of the universalizability test will depend on how the 
human formulates the maxim; if the human puts garbage into the test, the test will return garbage out.

It is likely that, regardless of the strengths of the human-in-the-loop model, fully automated AI 
agents will exist. Even if developing this kind of AI is irresponsible,
such developments are likely and will require ethics engines, or risk no consideration of ethics at all.
In such a world, the input parser in my ethics engine would have to be automated.
It is likely that, just as implementations of automated ethics choose 
a particular ethical theory and implement it, different implementations of such an input parser may 
adopt different interpretations of maxim formulation and morally relevant circumstances.

@@

My system's output could be converted into some actionable, useful response with an output parser, 
and then passed back to the AI agent. For example, if the AI agent is equipped to evaluate natural 
language prescriptions, the status of the maxim could be parsed into a natural language sentence. The 
input parser, categorical imperative, and output parser together constitute an ``ethics engine'' 
that AI developers could use in a variety of AI systems.

The ethics engine depicted above is a high-level example of one way to use my project to guide an artifical agent,
with additional work to parse the input and output of my implementation of the categorical imperative.
Effectively, the kind of automated ethics I implement could be part of a library that AI developers use to 
give AI agents the capacity for sophisticated ethical reasoning faithful to philosophical literature. 
This represents an improvement over existing automated ethics, which rarely captures the complexity 
of any ethical theory that philosophers plausibly defend.%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsection{Computational Ethics \label{computationalethics}%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
In addition to guiding AI agents, automated ethics can also help academic philosophers make philosophical
progress. Just as theorem provers make mathematics more efficient and push mathematicians to think 
precisely about the phenomena they model, computational ethics can help philosophers ask and answer
new philosophical questions. In Section \ref{joking}, I presented one example of the power of computational ethics
by using my system to isolate a necessary and sufficient condition for the act of lying, or knowingly saying
something false, to be wrong. In this section, I share another example of the kind of philosophical 
insight that computational ethics can prompt and analyze the value that this tool can offer to philosophers.\footnote{
I present a final example of computational ethics in Appendix \ref{weirdtests}, where I resolve an ambiguity 
in Korsgaard's argument for the wrongness of false promising using my system \citep{KorsgaardFUL}.} 
Fields from protein folding to game theory are uncovering new insights using computational tools, and computational
ethics harnesses this power to inspire similar progress in philosophy.%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsubsection{Example of a Philosophical Insight: Well-Formed Maxims%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
As presented in Section \ref{formalizingful}, in the process of developing my formalization of
the FUL, I discovered that certain kinds of maxims are badly formed, or inappropriate inputs to the 
universalizability test. The FUL is consistent only if it holds for ``well-formed maxims,''
such that neither the act nor goal are already achieved in the given circumstances. Precisely, 
a circumstance, act, goal tuple (c, a, g) is well-formed if $(\neg (c \longrightarrow a) ) \wedge 
(\neg(c \longrightarrow g))$, or if the circumstances do not imply the act or the goal. The insight
that the FUL does not apply to badly-formed maxims has philosophical value and serves as evidence 
of the potential of computational ethics. In the next section, I explain and explore this insight using 
Kantian literature. In the following section, I demonstrate the philosophical implications of well-formed
maxims by using this insight to resolve a tension between self-doubt and self-respect. The idea of well-formed
maxims can guide the formulation of maxims and thus may have implications for many parts of Kantian ethics.

@@

Kantians can make an even stronger claim about badly-formed maxims—because maxims are laws that you 
give to yourself, asking if you should will a maxim as you will it undermines your will's law-giving 
ability. The circumstances of a badly-formed maxim assume that the agent has willed the maxim. Under 
the Kantian acount of willing, willing a maxim is equivalent to giving the maxim to yourself as a law. 
When you will a maxim, you commit yourself to the maxim's end. You cannot simultaneously 
commit yourself to a maxim and ask if you should be committing to it. To will the maxim is to adopt it as 
law—so the question, ``should I be willing this?'' is paradoxical. Either you haven't actually made 
the maxim your law (and thus haven't yet committed to it), or you aren't actually asking 
the question (because the decision has already been made). Because a maxim is a law that you give to 
yourself, you cannot question it absent a sufficient reason, such as a change in the circumstances. 
To question a law arbitrarily is to not regard it as a law at all. This kind of questioning amounts to 
questioning the will's authority over itself, but this is impossible. The will definitionally has authority 
over itself, for that is what it is to be a will.

A skeptic may argue that we do often ask ``should I be doing this?'' as we do something. 
To understand this worry, I consider the maxim, 
``When dancing, I should just dance for the sake of dancing.'' While this maxim appears to be badly-formed (the 
circumstance ``dancing'' implies the act and goal of dancing), it is a question that practical reasoners 
do ask. I argue that the correct interpretation of this maxim is no longer a badly-formed maxim.

@@

Above, I defined badly formed maxims, a philosophical concept that I discovered
using computational ethics. In this section, I explore the implications of this concept for the ethical
tension between self-doubt and self-respect to show that computational ethics can result in insights with real 
philosophical weight. The debate between self-doubt and self-respect originates in epistemology, which 
values doubting your beliefs but also rationally requires that you believe that you are not mistaken 
(else you should update your beliefs). I present a parallel tension between self-doubt and self-respect
in ethics, where questioning your judgements is valuable but 
questioning a commitment as you make it is impossible. Ethical self-doubt and self-respect appear 
irresolvably opposed until they are understood through the lens of badly formed maxims. I argue that naive
conceptions of self-doubt are badly formed maxims in disguise. If we reformulate these maxims to be well-formed,
the tension between self-doubt and self-respect dissolves. I sketch the details of this argument in 
the rest of this section. I first introduce the tension between self-doubt and self-respect 
in epistemology, then explain the parallel tension in ethics, and finally present a resolution of this 
tension using badly-formed maxims. I conclude that well-formed maxims form a boundary condition for
the formulation of maxims, so they may resolve many debates in Kantian literature and ethics more generally.

@@

Computational ethics is at its infancy. Theorem provers are only now beginning to make headway in 
mathematics \citep{buzzardvideo}, even though theorem provers were first invented in the 1960's \citep{historyofITP}. 
In contrast, the first attempts to use theorem provers to automate deontic logic occurred in the last
few years. The fact that this nascent technology is already helping humans reach non-trivial ethical conclusions 
is reason to, at the very least, entertain the possibility of a future where computational ethics 
becomes as normal for philosophers as using a thought experiment.

To the skeptic, the fact that a theorem prover requires specialized knowledge outside of the field 
of philosophy indicates that the technology is nowhere near ready for universal use in philosophy 
departments. However, history indicates that as computing power increases and computer scientists make 
progress, computational ethics will become more usable. Theorem provers in mathematics began as toys 
incapable of proving that the real number 2 is not equal to the real number 1, but 
moving from basic algebra to Fields medal winning mathematics became possible in a
matter of years \citep{buzzardvideo}. Countless examples from the history of computer science, from the Turing 
Test to AI game playing to protein folding, demonstrate that progress in computer science can make seemingly 
obscure computer programs useful and usable in ways that exceed our expectations.
Programmable computers themselves initially began as unwieldy punch card readers, but their current ubiquity 
need not be stated. If computer scientists and philosophers invest in computational ethics, it could
become as commonplace in philosophy departments as reflective equilibrium. Just as computational tools
have amplified progress in healthcare and drug discovery, computational ethics has the potential to enable
great philosophical progress.%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsection{Automating Everyday Practical Reason \label{ordinarypeople}%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
In Sections \ref{AIethics} and \ref{computationalethics}, I outline how automated ethics can guide
artificial agents and philosophers respectively. This raises a natural question: can automated
ethics guide all human beings, not just academic philosophers, as we navigate the world and face ethical
dilemmas? Some may hope (or worry) that automated ethics could render everyday ethical reasoning obsolete. In 
this section, I argue that while computers should not replace human ethical reasoning entirely,
they can supplement and improve our ethical reasoning. I argue for a kind of human-computer symbiosis
in which computers offer ethical advice, arguments for particular moral judgements, and speed up moral
calculations without subverting human ethical reasoning entirely \citep{licklider}.

@@

In \emph{Universal Laws and Ends In Themselves}, O'Neill argues against the existence of an algorithm for
moral behavior. She points out that Kant draws an important distinction between a morally worthy maxim
and a morally worthy action: the latter requires a good will, or a will motivated by duty \citep[345]{oneilluniversallaws}. 
Moral behavior doesn't just require performing
a ``good'' action, but it requires acting on a morally worthy maxim from the motivation of duty, or 
doing the right thing because it is the right thing to do. It is this capacity 
for self-motivation that makes morality binding for rational beings; we must behave morally precisely 
because we have wills, or the ability to be motivated by ends
that we choose. Only rational beings have wills, so only rational beings can have good wills, so only 
rational beings can behave morally. Under this understanding
of moral behavior, it seems unlikely that a computer could behave morally
since a computer does not have motivation in the same way as a human being.\footnote{A parallel argument can also be made for virtue ethics. Virtuous
behavior requires not only a certain action, but also a certain disposition towards the action, so it seems
difficult for an AI agent to truly behave virtuously.}

The idea that a computer cannot behave morally does not preclude the kind 
of automated categorical imperative that I present in this thesis. O'Neill argues that the FUL
serves as a test of morally worthy maxims, and a implementation of an automated categorical imperative can 
identify this kind of maxim. Perhaps a computer cannot act on a morally relevant maxim from the motivation of duty, 
but it certainly can act on this maxim nonetheless. For example, a self-driving car can choose to swerve to hit a tree
to avoid injuring pedestrians in the crosswalk. This action may be one that acts on a morally worthy maxim
\emph{even if} the self-driving car is not motivated by duty. The discpline of machine ethics is
spurred by the recognition that, as automated agents become more powerful, they will need to make
morally consequential decisions. Automated agents may be incapable of moral motivation, but automated agents that mimic
moral behavior are better than agents that ignore morality entirely. AI agents are navigating a world 
inhabited by human beings, and their decisions impact us. Insofar as AI will operate in human society, 
the behavior of such AI should mimic the behavior of an ethical human being for our sakes, so that we 
can interact with it safely.

@@

Another challenge for automated Kantian ethics is that the FUL test requires that
a maxim be given as input. O'Neill notes that the test assumes ``that agents will have certain tentative 
plans, proposals and policies which they can consider, revise or reject or endorse and pursue'' \citep[343]{oneilluniversallaws}.
The FUL evaluates the moral worth of a maxim given as input, where this potential maxim is generated by 
the choices that an agent is faced with. As I argue in Section \ref{AIethics}, determining this potential maxim is a challenge for both human
and automated reasoners. Kant claims that determining an agent's potential maxim is so difficult because a maxim is their
prsonal understanding of their principle of action, and so we may never be able 
to know if the morally worthy action has been performed \cite[345]{oneilluniversallaws}. Reasoners are faced with
choices between potential actions and must determine the maxim, or principle, underlying each potential action.
This is equivalent to a ``mapping'' problem: agents are given situations or dilemmas as input and must map
these to maxims.

The challenge of mapping actions to maxims is a limitation of my system, but it is not insurmountable. In Section \ref{AIethics},
I argued that, before my system can be used in practice, it must be paired with an input parser that can
translate choices that an automated agent faces into maxims into a logic that my system can
evaluate. This need follows from the difficulty in mapping a potential action
to the maxim of action, whether concerning human action or machine action. As argued in Section \ref{AIethics},  
human-in-the-loop or heuristic-based approaches could resolve this issue. Determining the maxim of 
action is a challenge for Kantian human beings \citep{oneilluniversallaws}, so it is unsurprising that 
it is a major hurdle for automated Kantian agents. Overcoming this hurdle is not impossible, and progress
in automated ethics could address this concern.

@@

not already know what to do'' \citep[8]{tafani}.\footnote{Translated using Google Translate.} 

@@

\includegraphics[scale=0.47]{inputparser.png}

@@

could in turn be fed into the database of maxims to make the intput parser smarter. One example of 

@@

need for ethically intelligent AI agents. Tolmeijer et al. surveyed the state of the field of 

@@

%:%202=149%:%
%:%214=151%:%

@@

%:%233=163%:%
%:%245=165%:%

@@

%:%447=361%:%
%:%459=363%:%
%:%460=364%:%

@@

%:%533=430%:%
%:%545=432%:%
%:%546=433%:%

@@

%:%630=511%:%
%:%642=514%:%
%:%643=515%:%

@@

%:%764=635%:%
%:%765=636%:%

@@

%:%772=641%:%
%:%773=642%:%

@@

%:%796=660%:%
%:%808=662%:%
%:%809=663%:%

@@

%:%868=715%:%
%:%880=717%:%
%:%881=718%:%

@@

>>>>

First, I explore how my system can be used in practice to guide AI through moral dilemmas,
help academic philosophers make philosophical progress, and augment the everyday practical reasoning 
that we all perform as we navigate the world. I discuss these issues in Sections \ref{AIethics}, 
\ref{computationalethics}, and \ref{ordinarypeople} respectively and outline the further work
necessary to bring each of these applications to life. Much of this work has to do with how the input
maxim given to my system is formulated and the role of common sense or prior knowledge.

In Section \ref{amapossible}, I discuss theoretical objections to automated ethics, derived from the
common philosophical intuition that there is no algorithm for ethics. I explore concrete versions of 
this objection and make a theoretical argument for the possibility of the kind of automated ethics 
I implement in this thesis. Finally, in Section \ref{relatedwork}, I situate my project among related work.%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsection{Automated Moral Agents in Practice \label{AIethics}%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
In this section, I present the future work necessary for my system to guide AI 
in practice. As it stands, my system is a categorical imperative library that can
evaluate the moral status of a maxim. My project potentially 
serves as one component of an ``ethics engine'' that an AI system could use to make ethical decisions.
For example, when an AI system faces a moral dilemma in some internal representation, it could pass it 
to an input parser that translates the dilemma into a maxim in my logic. An output parser could then 
translate my system's output into a prescription for action that the AI system could act on. \begin{figure}
\includegraphics[scale=0.175]{AI_engine.jpeg}
\caption{An example of an ethics engine for artificial intelligence. This ethics engine passes a moral dilemma 
through an input parser, applies the automated categorical imperative test, and finally processes the 
output using an output parser, producing a prescription for action. I contribute the automated categorical 
imperative component.} \label{fig:AIengine}
\end{figure} Figure \ref{fig:AIengine} depicts the workflow of this example ethics engine. In order 
for my system to guide AI using this workflow, future work must develop such input and output parsers.

One of the biggest challenges for an ethics engine is the development of an input parser. An input parser 
for my implementation of automated Kantian ethics must translate a complex real-world situation into a 
circumstances, act, goal triple.
This requires that the input parser determine which circumstances are morally relevant
to a maxim, a controversial judgement. As introduced in Section \ref{whatisamaxim},
there is robust debate on the circumstances that should be considered when formulating a maxim, 
inspired by a common criticism of Kantian ethics called the tailoring objection. Recall that the 
tailoring objection is the worry that arbitrarily specific 
circumstances render any maxim universalizable. For example, consider the maxim ``When my name is Lavanya Singh 
and I am wearing a purple shirt and it is November 26th, I will lie in order to get some easy cash.'' 
Even if this maxim is willed universally, the circumstances are so 
specific that lying will not become the general mechanism for getting easy cash, so the lender will 
believe my lie and the maxim will remain effective. By tailoring the circumstances, any maxim can 
evade universalization.

In Section \ref{whatisamaxim}, I introduced the Kantian response to this criticism, which requires that 
the circumstances included in the formulation
of the maxim be morally relevant. In the example above, my purple shirt and the date have no bearing on 
the moral status of lying. On the other hand, consider the maxim, ``When I am unemployed, I will murder
someone in order to take their job.'' The circumstance ``when I am unemployed'' clearly has some moral
relevance to the murder in question; it speaks to the motivation for the murder.

@@

Translating everyday situations into appropriate maxims is the bulk of the reasoning that a Kantian 
human being does when making decisions, so it is unsurprising that formulating a maxim 
is one of the biggest obstacles to using my categorical imperative library
in an AI ethics engine. One solution is for a human being to perform the role of the input
parser by supervising the operation of an AI system. When the system stumbles onto
an ethical dilemma, the human could take over, formulate
the right question, and feed it into the categorical imperative library to see what action the categorical 
imperative would prescribe. Alternatively, AI developers can identify expected ethical dilemmas 
that the machine may face in advance and hardcode their own judgements for these dilemmas. 
For proponents of the ``human-in-the-loop'' model of AI ethics, in which
ethical AI requires that humans guide machines, this kind of human involvement may be a feature \citep{loop}.
Both of these solutions imply that the outcome of the universalizability test will depend on how the 
human formulates the maxim; if the human puts garbage into the test, the test will return garbage out.

It is likely that, regardless of the strengths of the human-in-the-loop model, fully automated AI 
will exist. Even if developing this kind of AI is irresponsible,
such developments are likely and will require ethics engines, or risk no consideration of ethics at all.
In such a world, the input parser in my ethics engine would have to be automated.
It is likely that, just as implementations of automated ethics choose 
a particular ethical theory and implement it, different implementations of such an input parser may 
adopt different interpretations of maxim formulation and morally relevant circumstances.

@@

My system's output could be converted into some actionable, useful response with an output parser, 
and then passed back to the AI system. For example, if the AI system is equipped to evaluate natural 
language prescriptions, the status of the maxim could be parsed into a natural language sentence. The 
input parser, categorical imperative, and output parser together constitute an ``ethics engine'' 
that AI developers could use in a variety of AI systems.

The ethics engine depicted above is a high-level example of one way to use my project to guide artifical intelligence,
with additional work to parse the input and output of my implementation of the categorical imperative.
The kind of automated ethics that I implement could be part of a library that AI developers use to 
give AI the capacity for sophisticated ethical reasoning faithful to philosophical literature. 
This represents an improvement over existing automated ethics, which rarely captures the complexity 
of any ethical theory that philosophers plausibly defend.%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsection{Computational Ethics \label{computationalethics}%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
In addition to guiding AI, automated ethics can also help academic philosophers make philosophical
progress. Just as theorem provers make mathematics more efficient and push mathematicians to think 
precisely about the phenomena they model, computational ethics can help philosophers ask and answer
new philosophical questions. In Section \ref{joking}, I presented one example of the power of computational ethics
by using my system to isolate a necessary and sufficient condition for lying, or knowingly saying
something false, to be wrong. In this section, I share another example of the kind of philosophical 
insight that computational ethics can prompt and analyze the value that this tool can offer to philosophers.\footnote{
I present a final example of computational ethics in Appendix \ref{weirdtests}, where I resolve an ambiguity 
in Korsgaard's argument for the wrongness of false promising using my system \citep{KorsgaardFUL}.} 
Fields from protein folding to game theory are uncovering new insights using computational tools, and computational
ethics harnesses this power to inspire similar progress in philosophy.%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsubsection{Example of a Philosophical Insight: Well-Formed Maxims%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
As presented in Section \ref{formalizingful}, in the process of formalizing
the FUL, I discovered that certain kinds of maxims are badly formed, or inappropriate inputs to the 
universalizability test. The FUL is consistent if it only holds for well-formed maxims,
such that neither the act nor goal are already achieved in the given circumstances. Precisely, 
a circumstance, act, goal tuple (c, a, g) is well-formed if $(\neg (c \longrightarrow a) ) \wedge 
(\neg(c \longrightarrow g))$, or if the circumstances imply neither the act nor the goal. The insight
that the FUL does not apply to badly-formed maxims has philosophical value and serves as evidence 
of the potential of computational ethics. In the next section, I explore this insight using 
Kantian literature. In the following section, I demonstrate the philosophical implications of well-formed
maxims by using this insight to resolve a tension between self-doubt and self-respect. The idea of well-formed
maxims can guide the formulation of maxims and thus may have implications for many parts of Kantian ethics.

@@

Kantians can make an even stronger claim about badly-formed maxims—because maxims are laws that you 
give to yourself, asking if you should will a maxim as you will it undermines your will's law-giving 
ability. The circumstances of a badly-formed maxim assume that the agent has willed the maxim. Under 
the Kantian acount of willing, willing a maxim is equivalent to giving the maxim to yourself as a law. 
When you will a maxim, you commit yourself to the maxim's end. You cannot simultaneously 
commit yourself to a maxim and ask if you should be committing to it. To will the maxim is to adopt it as 
law—so the question, ``should I be willing this?'' is paradoxical. Either you haven't actually made 
the maxim your law (and thus haven't yet committed to it), or you aren't actually asking 
the question (because the decision has already been made). Because a maxim is a law that you give to 
yourself, you cannot question it absent a sufficient reason, such as a change in the circumstances. 
To question a law arbitrarily is to not regard it as a law at all. This kind of questioning amounts to 
questioning the will's authority over itself, but this is impossible.\footnote{Recall from Section \ref{kantianethics} 
that \citet{velleman} presents this argument for the inescapability of practical reason.} The will 
definitionally has authority over itself, for that is what it is to be a will.

A skeptic may argue that we do often ask ``should I be doing this?'' as we do something. 
To understand this worry, I consider the maxim, 
``When dancing, I should just dance for the sake of dancing.'' While this maxim appears to be badly-formed (the 
circumstance ``dancing'' implies the act and goal of dancing), it is one that practical reasoners 
do consider. I argue that the correct interpretation of this maxim is no longer a badly-formed maxim.

@@

Above, I defined badly formed maxims, a philosophical concept that I discovered
using computational ethics. In this section, I explore the implications of this concept for the ethical
tension between self-doubt and self-respect to show that computational ethics can result in insights with real 
philosophical weight. The debate between self-doubt and self-respect originates in epistemology, which 
values questioning your beliefs but also rationally requires that you believe that you are not mistaken 
(else you should update your beliefs). I present a parallel tension between self-doubt and self-respect
in ethics, where questioning your judgements is valuable but 
questioning a commitment as you make it is impossible. Ethical self-doubt and self-respect appear 
irresolvably opposed until they are understood through the lens of badly formed maxims. I argue that naive
conceptions of self-doubt are badly formed maxims in disguise. If we reformulate these maxims to be well-formed,
the tension between self-doubt and self-respect dissolves. I sketch the details of this argument in 
the rest of this section. I first introduce the tension between self-doubt and self-respect 
in epistemology, then explain the parallel tension in ethics, and finally present a resolution of this 
tension using badly-formed maxims. I conclude that well-formed maxims form a boundary condition for
the formulation of maxims, so they may resolve many debates in Kantian literature and ethics more generally.

@@

Computational ethics is in its infancy. Theorem provers are only now beginning to make headway in 
mathematics \citep{buzzardvideo}, even though theorem provers were first invented in the 1960's \citep{historyofITP}. 
In contrast, the first attempts to use theorem provers to automate deontic logic occurred in the last
few years. The fact that this nascent technology is already helping humans reach non-trivial ethical conclusions 
is reason to, at the very least, entertain the possibility of a future where computational ethics 
becomes as normal for philosophers as using a thought experiment.

To the skeptic, the fact that a theorem prover requires specialized knowledge outside of the field 
of philosophy indicates that the technology is nowhere near ready for universal use in philosophy 
departments. However, history indicates that as computing power increases and computer scientists make 
progress, computational ethics will become more usable. Theorem provers in mathematics began as toys 
incapable of proving that the real number 2 is not equal to the real number 1, but 
moving from basic algebra to Fields medal winning mathematics became possible in a
matter of years \citep{buzzardvideo}. Countless examples from the history of computer science, from the Turing 
Test to AI game playing to protein folding, demonstrate that progress in computer science can make seemingly 
obscure computer programs useful and usable in ways that exceed our expectations.
Programmable computers themselves initially began as unwieldy punch card readers, but their current ubiquity 
need not be stated. If computer scientists and philosophers invest in computational ethics, it could
become as commonplace in philosophy departments as reflective equilibrium. Just as computational tools
have amplified progress in healthcare and drug discovery, computational ethics has the potential to enable
great philosophical progress.%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsection{Automating Everyday Practical Reason \label{ordinarypeople}%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
In Sections \ref{AIethics} and \ref{computationalethics}, I outline how automated ethics can guide
AI and philosophers respectively. This raises a natural question: can automated
ethics guide all human beings, not just academic philosophers, as we navigate the world and face ethical
dilemmas? Some may hope (or worry) that automated ethics could render everyday ethical reasoning obsolete. In 
this section, I argue that while computers should not replace human ethical reasoning entirely,
they can supplement and improve our ethical reasoning. I argue for a kind of human-computer symbiosis
in which computers offer ethical advice, arguments for particular moral judgements, and speed up moral
calculations without subverting human ethical reasoning entirely \citep{licklider}.

@@

In \emph{Universal Laws and Ends In Themselves}, O'Neill argues against the existence of an algorithm for
moral behavior. She points out that Kant draws an important distinction between a morally worthy maxim
and a morally worthy action: the latter requires a good will, or a will motivated by duty \citep[345]{oneilluniversallaws}. 
Moral behavior doesn't just require performing
a ``good'' action, but it requires acting on a morally worthy maxim from the motivation of duty, or 
doing the right thing because it is the right thing to do. It is this capacity 
for self-motivation that makes morality binding for rational beings; we must behave morally precisely 
because we have wills, or the ability to be motivated by ends
that we choose. Only rational beings have wills, so only rational beings can have good wills, so only 
rational beings can behave morally. Under this understanding
of moral behavior, it seems unlikely that a computer could behave morally
since a computer does not have motivation in the same way as a human being.\footnote{A parallel argument can also be made for virtue ethics. Virtuous
behavior requires not only a certain action, but also a certain disposition towards the action, so it seems
difficult for AI to truly behave virtuously.}

The idea that a computer cannot behave morally does not preclude the kind 
of automated categorical imperative that I present in this thesis. O'Neill argues that the FUL
serves as a test of morally worthy maxims, and a implementation of an automated categorical imperative can 
identify this kind of maxim. Perhaps a computer cannot act on a morally relevant maxim from the motivation of duty, 
but it certainly can act on this maxim nonetheless. For example, a self-driving car can choose to swerve to hit a tree
to avoid injuring pedestrians in the crosswalk. This action may be one that acts on a morally worthy maxim
\emph{even if} the self-driving car is not motivated by duty. The discpline of machine ethics is
spurred by the recognition that, as automated agents become more powerful, they will need to make
morally consequential decisions. Automated agents may be incapable of moral motivation, but automated agents that mimic
moral behavior are better than agents that ignore morality entirely. AI is navigating a world 
inhabited by human beings, and its decisions impact us. Insofar as AI will operate in human society, 
the behavior of such AI should mimic the behavior of an ethical human being for our sakes, so that we 
can interact with it safely.

@@

Another challenge for automated Kantian ethics is that the FUL test requires that
a maxim be given as input. O'Neill notes that the test assumes ``that agents will have certain tentative 
plans, proposals and policies which they can consider, revise or reject or endorse and pursue'' \citep[343]{oneilluniversallaws}.
The FUL evaluates the moral worth of a maxim given as input, where this potential maxim is generated by 
the choices that an agent is faced with. As I argue in Section \ref{AIethics}, determining this potential maxim is a challenge for both human
and automated reasoners. Kant claims that determining an agent's potential maxim is difficult because a maxim is their
personal understanding of their principle of action, so we may never be able 
to know if the morally worthy action has been performed \cite[345]{oneilluniversallaws}. Reasoners are faced with
choices between potential actions and must determine the maxim, or principle, underlying each potential action.
This is equivalent to a ``mapping'' problem: agents are given situations or dilemmas as input and must map
these to maxims.

The challenge of mapping actions to maxims is a limitation of my system, but it is not insurmountable. In Section \ref{AIethics},
I argued that, before my system can be used in practice, it must be paired with an input parser that can
translate choices that an AI system faces into maxims into a logic that my system can
evaluate. This need follows from the difficulty in mapping a potential action
to the maxim of action, whether concerning human action or machine action. As argued in Section \ref{AIethics},  
human-in-the-loop or heuristic-based approaches could resolve this issue. Determining the maxim of 
action is a challenge for Kantian human beings \citep{oneilluniversallaws}, so it is unsurprising that 
it is a major hurdle for automated Kantian agents. Overcoming this hurdle is not impossible, and progress
in automated ethics could address this concern.

@@

not already know what to do'' \citep[8]{tafani}.\footnote{Translated from Italian to English using Google Translate.} 

@@

\includegraphics[scale=0.19]{inputparser.jpeg}

@@

could in turn be fed into the database of maxims to make the input parser smarter. One example of 

@@

need for ethically intelligent AI. Tolmeijer et al. surveyed the state of the field of 

@@

%:%194=148%:%
%:%203=150%:%

@@

%:%225=162%:%
%:%234=164%:%

@@

%:%439=359%:%
%:%440=360%:%
%:%449=363%:%

@@

%:%525=429%:%
%:%526=430%:%
%:%535=432%:%

@@

%:%622=509%:%
%:%623=510%:%
%:%632=513%:%

@@

%:%762=634%:%
%:%763=635%:%

@@

%:%770=641%:%
%:%771=642%:%

@@

%:%788=657%:%
%:%789=658%:%
%:%798=662%:%

@@

%:%860=714%:%
%:%861=715%:%
%:%870=717%:%

@@

%:%919=756%:%
%:%920=757%:%
